A general and efficient representation of Ancestral Recombination Graphs

### 2022-09-01 paper.tex
Paragraph on downstream apps for full ARGs

>>>>

% Do downstream applications actually make use of such full ARGs?
Downstream applications using estimated ARGs to date rarely
make use of such complete recombination information.
The \relate\ selection statistic~\citep{speidel2019method}
computes a $p$-value for selection [FIXME] for each
local tree independently by comparing the rates of coalescence
at various time-points. For cases like strong positive selection,
where we expect there to be a clear signal in the local tree,
such tree-by-tree methods can be simple and powerful.
In their method
for estimating dispersal rates and the locations of genetic
ancestors,
\citep{osmond2021estimating} downsample trees along the genome
so that they can be regarded as approximately independent.
The SIA method for detecting selection~\citep{hejase2022deep}
encodes local trees as a set of lineage counts at discrete
time intervals, and uses these as feature for a
type of machine learning algorithm
that takes ``temporal'' correlations into account. Thus,
while SIA does use some of the information about local tree correlation,
clearly much of the detail about recombination
events in a ``full'' ARG is lost.
The \texttt{tsdate} algorithm (and related naive estimator
for ancestral location) uses much more between-tree
information~\citep{wohns2022unified}, explicitly using the gARG
encoding to reason about ancestral nodes.
A major strength of this approach is that it can be used on
any gARG, ranging from estimates containing relatively little
information about recombination (e.g., as estimated by \relate)
to ``full'' ARGs containing complete recombination information
(e.g., as estimated by \kwarg), and will fully utilise
any between-tree correlation information that is present.
